The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-10-08 Paper.qmd
Added barplot for caregiver response to last child signal.
New stuff in dataprep and regularize. Minor changes caused by the visual editor.

|  |  |
|--------------------------|----------------------------------------------|
| Variable | Survey question wording |
| Crying Frequency | How often does your child cry? |
| Sad Frequency | How often is your child sad? |
| Tantrum Frequency | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency | How often does your child leave the house without telling you when they get mad? |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child? |
| Child Alloparenting Frequency | How often does your child take care of their other siblings? |

@@

|  |  |
|--------------------------|----------------------------------------------|
| Variable | Survey question wording |
| Relative Need | How much time, effort, and resources does this child need compared to your other children? |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

@@

::: {#fig-signal_corrplot}
```{r}
signal_corrplot
```
Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
:::

@@

## Signaling behavior is negatively associated with caregiver education
[delete this section if we remove conflict from the main model? or mention on significant when controlling for conflict]

@@

::::::::::: {#fig-effect_allo_by_sex}
```{r}
signaling_data_plot_allo_by_sex
```

@@

[add sad freq2]
Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children. Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor. The end result was `{r} rel_need_table[1]` less needy children, `{r} rel_need_table[2]` children who were as needy as their siblings, and `{r} rel_need_table[3]` more needy children.

@@

[do we want to mentioned distributions of the actual data?]
Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NegativeResponseBinary > 0)` negative response). However, they were also more likely to respond in a neutral or negative manner rather than a positive one when caretakers believed signaling came in response to the child being punished (`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NegativeResponseBinary > 0)` negative responses).

@@

|  |  |
|--------------------------|----------------------------------------------|
| Variable | Survey question wording |
| IllnessAvoidance1 | If an illness is 'going around', my child will get it. |
| IllnessAvoidance2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| IllnessAvoidance3 | My child’s immune system protects him/her from most illnesses that other kids get. |

@@

::: {#stbl-need-investment}
```{r}
modelsummary(polrModels$Model$Invest, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

>>>>

|  |  |
|----|----|
| Variable | Survey question wording |
| Crying Frequency | How often does your child cry? |
| Sad Frequency | How often is your child sad? |
| Tantrum Frequency | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency | How often does your child leave the house without telling you when they get mad? |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child? |
| Child Alloparenting Frequency | How often does your child take care of their other siblings? |

@@

|  |  |
|----|----|
| Variable | Survey question wording |
| Relative Need | How much time, effort, and resources does this child need compared to your other children? |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

@@

::: {#fig-signal_corrplot}
```{r}
signal_corrplot
```

Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
:::

@@

## Signaling behavior is negatively associated with caregiver education

\[delete this section if we remove conflict from the main model? or mention on significant when controlling for conflict\]

@@

::: {#fig-effect_allo_by_sex}
```{r}
signaling_data_plot_allo_by_sex
```

@@

\[add sad freq2\] Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children. Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor. The end result was `{r} rel_need_table[1]` less needy children, `{r} rel_need_table[2]` children who were as needy as their siblings, and `{r} rel_need_table[3]` more needy children.

@@

\[do we want to mentioned distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NegativeResponseBinary > 0)` negative response). However, they were also more likely to respond in a neutral or negative manner rather than a positive one when caretakers believed signaling came in response to the child being punished (`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NegativeResponseBinary > 0)` negative responses).

@@

|  |  |
|----|----|
| Variable | Survey question wording |
| IllnessAvoidance1 | If an illness is 'going around', my child will get it. |
| IllnessAvoidance2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| IllnessAvoidance3 | My child’s immune system protects him/her from most illnesses that other kids get. |

@@

::: {#stbl-need-investment}
```{r}
modelsummary(polrModels$Model$Invest, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```
:::
